---
title: "MV_euc"
output:
  html_document: default
  pdf_document: default
date: "2024-04-06"
---

Load package and set-up directory 
```{r setup, include=FALSE}
#install.packages('ggdendro')
#install.packages('ggheatmap')
#install.packages('dendextend')


knitr::opts_knit$set(root.dir = "~/stats/euc")
library(phyloseq)
library(vegan)
library(ggplot2)
library(ggpubr)
library(ggrepel)
library(glue)
library(cluster)
library(dplyr)
library(scales)
library(tidyr)
library(viridis)

library(ggVennDiagram)
library(plotly)
library(VennDiagram)

library(GGally)
library(ggdendro)
library(ggheatmap)
library(dendextend)

#library(ggforce)
``` 

#load data
```{r}
asv=read.table( "~/stats/median_asv_MV_euc.csv", sep=",", row.names=1, header=TRUE, check.names=FALSE)
meta=read.table("~/stats/median_meta_MV_euc.csv", sep=",", row.names=1, header=TRUE, stringsAsFactors = TRUE)
 #change time for exp. 2 from 22-45 to 1-24
 meta$time <- ifelse(meta$experience == 'E2', meta$time - 21, meta$time)

 meta <- meta %>% select(-c(X, X.1, X.2, X.3, X.4))

 taxa=read.table("~/stats/median_taxa_MV_euc.csv", sep=",", row.names=1, header=TRUE)

```

#check for colinearity between dic, doc, time and bioreactor for each experiment
```{r}

#subset meta based on exp
meta_E1 <- subset(meta, meta$experience == 'E1' & short_sample_type == 'R')

meta_E2 <- subset(meta, meta$experience == 'E2' & short_sample_type == 'W')


ggpairs(meta_E1, columns = c('dic_raw', 'doc_raw', 'time', 'ph_raw', 'do_percent_raw', 'bioreactor'), axisLabels = 'show', ggplot2::aes(colour = bioreactor, ))

ggpairs(meta_E2, columns = c('dic_raw', 'doc_raw', 'time', 'ph_raw', 'do_percent_raw', 'bioreactor'), ggplot2::aes(colour = bioreactor))

sum_E1 <- summary(meta_E1)
aov(meta_E1$dic_raw ~ meta_E1$bioreactor)
write.csv(sum_E1, '~/stats/sum_E1.csv')

sum_E2 <- summary(meta_E2)
write.csv(sum_E2, '~/stats/sum_E2.csv')

#DOC and time are strongly correlated for exp. 2 so DOC won't be included in the dbRDA
```


NMDS all 
```{r} 

#remove all ASV that are not in any sample
asv <- asv[rowSums(asv) > 0, ]

# asv_hellinger <- decostand(asv, method="hellinger") #only if want pcoa on hellinger instead of bray

meta_pcoa <- subset(meta, experience_short_sample_type == 'E1_R' | experience_short_sample_type == 'E2_W')
# 
ps <- phyloseq(
    otu_table(asv, taxa_are_rows = TRUE), #change to asv_hellinger if want hellinger
    tax_table(as.matrix(taxa)),
    sample_data(meta_pcoa)
  )

  # Perform PCoA using the "bray" dissimilarity metric
  ordu <- ordinate(ps, 'NMDS', 'bray')
    p <- plot_ordination(ps, ordu, color = "time", shape = "experience_short_sample_type", title = paste('NMDS all euc samples', '      ', 'stress =', signif(ordu$stress, 3), sep = '' )) +
       theme_minimal() +
  theme(
   panel.grid = element_blank(),
    text = element_text(size = 12),
    legend.position = "bottom",
    plot.title = element_text(hjust = 0.5),
    axis.text = element_text(color = "black") # Set axis text color to black
  ) +
    scale_fill_viridis_c(alpha = 1, begin = 0, end = 1, direction = 1, option = "D", aesthetics = "colour")

print(p)
  
  
    # Save the plot as a separate image file
  ggsave(("~/stats/NMDS/euc/NMDS_euc_all.pdf"), p, width = 8, height = 6, dpi = 300)
```

# NMDS for loop for different sample_types and experiment 
```{r} 

# Create an empty list to store the plots
plot_list <- list()
#type_exp == 'E2_W'
# Loop through unique values of "type_exp"
for (type_exp in unique(meta$experience_short_sample_type)) {
#type_exp = 'E1_R'
    # Subset the data for the current "type_exp"
  meta_sample <- subset(meta, experience_short_sample_type == type_exp)
  
  # Create a phyloseq object using the subsetted data
#asv_hellinger <- decostand(asv, method="hellinger") #only if want pcoa on hellinger instead of bray  
  ps <- phyloseq(
    otu_table(asv, taxa_are_rows = TRUE),
    tax_table(as.matrix(taxa)),
    sample_data(meta_sample)
  )
# 
#   
    # Perform NMDS using the "bray" dissimilarity metric
  ordu <- ordinate(ps, 'NMDS', 'bray')
  
  # Create the NMDS plot
  p <- plot_ordination(ps, ordu, color = "time", shape = "bioreactor"#, title = paste('NMDS ', type_exp, ' euc', '        ', 'stress =', signif(ordu$stress, 3), sep = '' )
                       )+
 theme_minimal() +
  theme(
   panel.grid = element_blank(),
    text = element_text(size = 12),
    plot.title = element_text(hjust = 0.5),
    axis.text = element_text(color = "black") # Set axis text color to black
  ) +
    scale_fill_viridis_c(alpha = 1, begin = 0, end = 1, direction = 1, option = "D", aesthetics = "colour")
  

  # Save the plot in the plot_list with a name starting with "PCoA_euc_"
  plot_list[[paste0("NMDS_euc_", type_exp)]] <- p
  # 
  print(p)
  # 
  # Save the plot as a separate image file
  filename <- paste0("NMDS_euc_", type_exp, ".pdf")
  ggsave(paste0("~/stats/NMDS/euc/NMDS_euc_", type_exp, ".pdf"), p, width = 8, height = 6, dpi = 300)
}

```


Prep data for db-RDA
```{r}
# Create a new column using the first two terms of the sample's name
asv_db <- as.data.frame(t(asv))
asv_db$experience_short_sample_type <- (substr(rownames(asv_db), 1, 4))
asv_db$experience_short_sample_type <- gsub('-', '_', asv_db$experience_short_sample_type)
asv_db$experience_short_sample_type <- as.factor(asv_db$experience_short_sample_type)
asv_db <- as.data.frame(asv_db) %>%
dplyr::select(experience_short_sample_type, everything())  # Reorder columns
```

#db-RDA on all to confirm experiment and sample_type are significantly differents 
```{r}
asv_all <- as.data.frame(t(asv))

 beta_dist_all <- vegdist(asv_all, method = "bray")
dbRDA_all <- capscale(beta_dist_all ~ Sample_type*experience + bioreactor, data = meta, na.action = na.omit)
anova_all <- anova(dbRDA_all, by = 'onedf', permu = 1000)

anova_all
```



#perform and plot db-RDA and varpart
```{r}
meta$position_pastille <- as.factor(meta$position_pastille)

i = 0
dbRDA_exp_type = list()  # Initialize an empty list before the loop

for (type_exp in unique(meta$experience_short_sample_type)) {
 i = i + 1

  # Subset the dataframe based on the current type_exp
  subset_asv <- asv_db[asv_db$experience_short_sample_type == type_exp, ]
  subset_asv <- subset_asv[, -which(names(asv_db) == 'experience_short_sample_type')]
  subset_meta <- meta[meta$experience_short_sample_type == type_exp, ]
  beta_dist <- vegdist(subset_asv, method = "bray")
  

  # Perform db-RDA
if (type_exp == 'E2_W') {
  dbRDA <- capscale(beta_dist ~ dic + time + bioreactor, data = subset_meta, na.action = na.omit)
} else {
  dbRDA <- capscale(beta_dist ~ dic + doc + dic:doc + time + bioreactor, data = subset_meta, na.action = na.omit)
}  
  #    
  dic=subset_meta$dic
  doc=subset_meta$doc
  
  
# Extract the site scores from dbRDA result
site_scores <- scores(dbRDA, display = "sites")
site_scores_df <- as.data.frame(site_scores)

# Extract the variable scores from dbRDA result
variable_scores <- scores(dbRDA, display = "bp")
variable_scores_df <- as.data.frame(variable_scores)

# Get the dimension names from the site scores
dim_names <- colnames(site_scores_df)

# Calculate the eigenvalues of the dbRDA result
eigenvals <- eigenvals(dbRDA)

# Calculate the percentage of variance explained by each dimension
var_explained <- eigenvals / sum(eigenvals) * 100

# Create a dataframe combining site scores and metadata
plot_data <- cbind.data.frame(site_scores_df, subset_meta)

  
  
dbrdarock <- ggplot() +
  geom_hline(yintercept = 0, linetype = "solid", color = "gray") +
  geom_vline(xintercept = 0, linetype = "solid", color = "gray") +
  geom_point(data = plot_data, aes(x = CAP1, y = CAP2, shape = bioreactor, color = time), size = 2) +
  geom_segment(data = variable_scores_df, aes(x = 0, y = 0, xend = CAP1, yend = CAP2), 
               arrow = arrow(length = unit(0.15, "cm")), color = "black") +
  
  # Add text labels for environmental variable names
  geom_text(
    data = variable_scores_df,
    aes(x = CAP1, y = CAP2, label = rownames(variable_scores_df)),
    vjust = -1, hjust = 0.5,
    color = "#c45015"
  ) +

  # scale_color_manual(values = c("forestgreen","goldenrod3","firebrick3", "deepskyblue3" ), name = "bioreactor",) + 
  xlab(paste(dim_names[1], " (", round(var_explained[1], 2), "%)")) + #naming the x axis
  ylab(paste(dim_names[2], " (", round(var_explained[2], 2), "%)")) + #naming the y axis
#  geom_text(data = plot_data, aes(x = CAP1, y = CAP2, label = sample_name), hjust = 0.5, vjust = -0.5, size = 3.5) + #adding the sample name
  theme_minimal() +
  theme(text = element_text(size = 15))

# Store the results in the list
dbRDA_exp_type[[i]] <- list(
  dbRDA = dbRDA,
  plot = dbrdarock
)
print(dbrdarock)
#figure out how to save if want to publish, if not, it's below the script #maybe use ggVennDiagram instead?
#variance partitioning for every dbRDA #varpart only allows for 4 explanatory variables so the interaction has been removed from the analysis
if (type_exp == 'E2_W') { #unsure if it's the proper type_exp
 varpart_full <- varpart(beta_dist, ~ dic, ~ time,~ bioreactor, data = subset_meta)
  Venn_varpart <-  plot(varpart_full, bg = c("hotpink","skyblue"), Xnames = c('DIC', 'Time', 'Bioreactor'))
} else {
  varpart_full <- varpart(beta_dist, ~ dic,~ doc,~ time,~ bioreactor, data = subset_meta)
     Venn_varpart <-  plot(varpart_full, bg = c("hotpink","skyblue"), Xnames = c('DIC', 'DOC', 'Time', 'Bioreactor'))
}  

  # Set path for saving plots
plot_path="~/stats/dbRDA/euc/"

  #save plots
name <- paste('median_', type_exp, "_dbRDA_results_", i, ".jpeg", sep = "")
ggsave(file.path(plot_path, name), width = 6, height = 4, plot = dbrdarock)
}

```

#statistical tests db-RDA
```{r}

meta$experience_short_sample_type <- as.factor(meta$experience_short_sample_type)
meta$bioreactor <- as.factor(meta$bioreactor)
meta$experience <- as.factor(meta$experience)

i = 0
# Loop through unique values of "experience_short_sample_type"
for (type_exp in unique(meta$experience_short_sample_type)) {
  
  i = i + 1

  # Now you can access the dbRDA result and plot as follows:
  anova_exp_results <- anova(dbRDA_exp_type[[i]][[1]], by="onedf", permu=1000)

  Rsquare_exp <- RsquareAdj(dbRDA_exp_type[[i]][[1]])

  #summary results dbRDA
  dbRDA_exp <- summary(dbRDA_exp_type[[i]][[1]])
   
  
  # Set path for saving plots
   plot_path = "~/stats/dbRDA/euc/anova_summary"

   # Save files for anova
   name <- paste(type_exp, "_median_anova_results_", i, ".csv", sep = "")
   write.csv(anova_exp_results, file = file.path(plot_path, name))
   #save file for
   name_R <- paste(type_exp, "_median__Rsquare_results_", i, ".csv", sep = "")
     write.csv(Rsquare_exp, file = file.path(plot_path, name_R))

}
```

Prep data for dendrogram
```{r}
# Create a new column using the first two terms of the sample's name
asv_dend <- as.data.frame(t(asv))
asv_dend$experience_short_sample_type_bio <- paste((substr(rownames(asv_dend), 1, 4)), (substr(rownames(asv_dend), 10, 11)), sep = '_')
asv_dend$experience_short_sample_type_bio <- gsub('-', '_', asv_dend$experience_short_sample_type_bio)
asv_dend$experience_short_sample_type_bio <- as.factor(asv_dend$experience_short_sample_type_bio)
asv_dend <- as.data.frame(asv_dend) %>%
dplyr::select(experience_short_sample_type_bio, everything())  # Reorder columns

asv_dend$name <- rownames(asv_dend)
asv_dend$time <- as.integer(substr(asv_dend$name, 7, 8))
asv_dend$time <- ifelse(grepl('E2_W', asv_dend$experience_short_sample_type_bio), asv_dend$time - 21, asv_dend$time)

asv_dend$time <- ifelse(asv_dend$time < 10, paste('0', asv_dend$time, sep = ''), asv_dend$time)
asv_dend$time <- paste('T', asv_dend$time, sep  = '')
```

forloop of dendrograms for every bioreactors and every day
```{r}

i = 0
# Loop through unique values of "experience_short_sample_type"
for (den in unique(asv_dend$experience_short_sample_type_bio)) {
i = i + 1
 
  if (grepl('E1_W', den) | grepl('E2_R', den)) {
    next  
  }

#create the asv DF only containing the experiments of interest
sub_asv <- subset(asv_dend, experience_short_sample_type_bio == den)

#use time as the rowname for better labelling in the graphs
rownames(sub_asv) <- sub_asv$time

#remove the column 'experience_short_sample_type_bio'
sub_asv <- sub_asv %>% dplyr::select(-c(experience_short_sample_type_bio, time, name))


veg <- vegdist(sub_asv)

hcl <- hclust(veg)

plot_test <- as.dendrogram(hcl)
plot(plot_test, ylab = "Dissimilarity", xlab = 'time', ylim = c(0, 1), main = den)

}
```
