
# Code 
## Installing librarie
R version 4.3.1 
```{r, eval=FALSE}
# install.packages("devtools")
# devtools::install_github("andyofsmeg/ggTag")
# install.packages("forcats")
# install.packages("dplyr")
# if (!require("BiocManager", quietly = TRUE))
#     install.packages("BiocManager")
# BiocManager::install(version = "3.17")
# BiocManager::install('phyloseq')
# install.packages("randomcoloR")
# install.packages("ggplot2")
# install.packages("tidyr")
# install.packages("stringr")
# install.packages("ggtext")
# install.packages("DT")
```
## Loading the libraries
```{r, message=FALSE, warning=FALSE}
library(devtools)
library(ggTag) # To extract title from plots
library(forcats) # To reorder our factor (function "fct_relevel")
library(dplyr) # Dataframe manipulation (notably function "pull")
library(phyloseq) # Very pratical library for the analysis of amplicon data 
library(randomcoloR) # Generate sets of random colors
library(ggplot2) # Generate plot 
library(tidyr) # Dataframe manipulation (function "separate")
library(stringr) # Makes working with strings as easy as possible (function "str_replace")
library(ggtext) # Allow the use of markdown text (used to make names italic)
library(DT)# Make beautiful table 
```


```{r}
asv=read.table("~/median/median_bac.csv", sep=",", row.names=1, header=TRUE, check.names=FALSE)
taxa=read.table("~/median/median_taxa_bac.csv", sep=",", row.names=1, header=TRUE)
meta=read.table('~/median/median_meta_bac.csv', sep=",", row.names=1, header=TRUE)

```

Remove unwanted sample and column from the sample metadata file 
```{r, echo=FALSE}
#meta=subset(meta, !Sample_type=='Peatbog') # Remove sample from the peatbog
meta=subset(meta, select=-c(is.neg))
meta$position_pastille[meta$Sample_type == "Groundwater"] <- NA
```


```{r}
library(tidyr)
meta$sample_name=row.names(meta) # Create new column with row.name
# Separate new column based on hyphen
meta=separate(meta,
              col=sample_name,
              into=c("experience", "short_sample_type", "time", "bioreactor", "replicate"),
              sep="-")
meta$sample_name=row.names(meta) # reset row names because R

meta$experience=gsub("E", "Experience ", meta$experience) # In the column experience, replace E with Experience
meta$bioreactor=gsub("B0","Aquifer", meta$bioreactor) # In the column bioreactor, replace B0 with Aquifer
meta$bioreactor=gsub("B", "Bioreactor ",meta$bioreactor) # In the column bioreactor, replace B with Bioreactor

```


substract 21 from the days of exp. 2
```{r}
meta$time <- as.numeric(gsub('T', '', meta$time))
meta$time <- ifelse(meta$experience == 'Experience 2' & meta$time != '0', meta$time - 21, meta$time)
meta$time <- ifelse(meta$time < 10, paste(0, meta$time, sep = ''), meta$time)
#meta$time <- order(meta$time)
meta$time <- paste('T', meta$time, sep = '')

```

```{r}
 # Creating new column for our iterator by combining experience # and sample type
 meta$experience_sample=paste(meta$experience,"-",meta$Sample_type)

 # create the new column to be used for the x axis (representing both times and replicate #)
 meta$time_replicate=with(meta, ifelse(bioreactor=="Aquifer",paste("T00.", meta$replicate, sep = ""),
                                       ifelse(experience=="Experience 2" & Sample_type=="Rock",
                                       paste(meta$time,paste("R.", meta$replicate), sep = "-"),
                                       meta$time)))

 # While we are here let's also get rid of columns we won't need
 meta=subset(meta, select=-c(position_pastille, temp, ph, do_percent, dic, doc, short_sample_type, X, X.1, X.2, X.3, X.4))
```


Generate long format dataframe for plotting 
```{r}
library(phyloseq)
# Merge into phyloseq object
ps=phyloseq(otu_table(asv, taxa_are_rows=TRUE), tax_table(as.matrix(taxa)), sample_data(meta))
# Select taxonomic rank
#
#
taxa_rank="Family"
#
#
#
ps_rel_abund=transform_sample_counts(ps, function(x) x/sum(x)) # get abundance in %
glom=tax_glom(ps_rel_abund, taxrank = taxa_rank) # agglomerate taxa at the desired taxonomic level

# Concatenate experience and short_sample_type into new column name
# This will allow me to loop over this new variable in order to get a more accurate relative abundance of the taxa
df=psmelt(glom)
```


## Identify most abundant taxa and assign color 

Create a list of unique most abundant taxas in each sample
```{r, eval=FALSE}
library(dplyr)
list_of_all_taxonomic_rank= list() # Create empty list
i = 0
# Loop over every sample to combine taxa that are not the top 5 in new group called Others
for (each_sample in unique(df$Sample)){
    i=i+1
    sample=subset(df, Sample==each_sample)
    total_abundance=aggregate(sample$Abundance, by=list(taxa_rank=sample[[taxa_rank]]), FUN=sum)
    top=head(total_abundance[order(total_abundance$x, decreasing= T),], n = 5)
    others_df=sample[!sample[[taxa_rank]] %in% top$taxa_rank,] # identify other
    others_list=pull(others_df, taxa_rank)
    sample[sample[[taxa_rank]]%in% others_list,][[taxa_rank]]="Others"
    list_of_all_taxonomic_rank[[i]]=sample
}
df2=do.call("rbind",list_of_all_taxonomic_rank) # combine all generated list into one dataframe
unique_taxon=data.frame(unique(df2[[taxa_rank]]))
name=colnames(unique_taxon)
names(unique_taxon)[names(unique_taxon)==name]=as.character(taxa_rank) # Rename column

#rename the aquifer to be included with E1_R 
df2 <- df2 %>%
  mutate(experience_sample = ifelse(bioreactor == "Aquifer" & experience_sample == "Experience 1 - Groundwater", 
                                    "Experience 1 - Rock", 
                                    experience_sample))

```

## Create new color range (only needed once)
```{r, eval=FALSE}
library(randomcoloR)
library(stringr)
library(forcats)
library(ggplot2)
n=nrow(unique_taxon)
palette=distinctColorPalette(n, runTsne = TRUE)
unique_taxon[[taxa_rank]]=factor(unique_taxon[[taxa_rank]])
names(palette)=levels(unique_taxon[[taxa_rank]])
palette[["Others"]]="#E1E1E1"

#  recreate palette with markdown to italize name and remove the underscore after Unclassified
all_names=data.frame(names(palette))
names_markdown=all_names %>%
  mutate(names.palette.=str_replace(names.palette., "(.*)","*\\1*"),
         names.palette.=str_replace(names.palette., "\\*Unclassified_(.*)\\*","Unclassified *\\1*"),
         names.palette.=str_replace(names.palette., "\\*Others\\*", "Others"))
list_names=as.vector(names_markdown$names.palette.)
# Replace names of obj
names(palette)=c(list_names)

# Making the same modification to the taxa_rank names from the legend to the taxa_rank names in the dataframe
df2[[taxa_rank]]=str_replace(df2[[taxa_rank]], "(.*)","*\\1*")
df2[[taxa_rank]]=str_replace(df2[[taxa_rank]], "\\*Unclassified_(.*)\\*","Unclassified *\\1*")
df2[[taxa_rank]]=str_replace(df2[[taxa_rank]], "\\*Others\\*", "Others")


# Ordering the legend in alphabetical order
legend_raw=unique(df2[[taxa_rank]]) #Extract legend as text
ordered_legend=sort(legend_raw) # order alphabetically
reordered_legend=fct_relevel(ordered_legend, "Others") # move "Others" to the beginning
final_legend=levels(reordered_legend) # Extract the levels in a new object

my_scale <- scale_fill_manual(name=as.character(taxa_rank), breaks=paste(final_legend), values = palette, na.translate=FALSE, drop=TRUE, limits = force) # Recreate color fill
```

Saving and reloading color range 
```{r, eval=FALSE}
saveRDS(my_scale,paste('~/stats/histogramme/bac/top_', taxa_rank, '_color_palette.rds', sep = '')) # saving color palette
my_scale=readRDS(paste('~/stats/histogramme/bac/top_', taxa_rank, '_color_palette.rds', sep = '')) # Load color palette
```

## Plotting your figures
```{r, eval=FALSE}
#Loop for plotting results
library(forcats)
library(ggtext)
plot_list = list()
i = 0
for (habitat in unique(df2$experience_sample)){
  i = i + 1
  df_sample=subset(df2, experience_sample==habitat)
  
  p=ggplot(df_sample, aes(x=time_replicate, weight=Abundance, fill=fct_reorder(.data[[taxa_rank]],Abundance,.desc=FALSE))) + # .data is very important to force the evaluation of the input variables (taxonomic_rank)
  geom_bar() +
  labs(y ='Relative abundance (%)', x="Time") +
  scale_y_continuous(expand = c(0,0)) +
  theme_classic() +
  theme(text = element_text(size = 10),
        axis.title=element_text(size=8),
        legend.title=element_text(size=10),
        plot.title = element_text(hjust =0.5),
        axis.text.x = element_text(angle=70,vjust = 0.5, size=7),
        legend.position="bottom", #change to right
        legend.text=element_markdown(size=10),
        legend.key.size = unit(0.5, 'cm'), # adjusting the size of the legend colored box
        strip.background = element_blank()) + # remove facet_grid box background
  ggtitle(paste(habitat)) +
  my_scale +
  facet_grid(~bioreactor, scales="free", space="free") +
  guides(fill=guide_legend(ncol = 5, title=taxa_rank, title.position="top", title.hjust = 0.5, reverse=FALSE)) #change nrow = 6 to ncol = 1
  plot_list[[i]] = p
}
# plot_list

# Set path for saving plots
plot_path="~/stats/histogramme/bac/"
for (i in plot_list[]){
  title=extractGGTitle(i) # extract title from plot
  titlepdf=paste('median_', taxa_rank,"-", title,".pdf", sep="") # add.pdf to the end of every title to save figure as pdf
  ggsave(as.character(titlepdf),plot=i, device="pdf", path=plot_path, width=27.94, height = 21.59, units="cm") # save plot
}
```

# Export tables
```{r, eval=FALSE}
df2$"Relative abundance"=(paste( df2$experience_sample, "_", df2$time_replicate,"_", df2[[taxa_rank]]))
supplementary_dataframe=aggregate(x=df2$Abundance, by=list(df2$"Relative abundance"), FUN=sum)

# Save tables
title_df=paste("~/stats/histogramme/bac/median_", taxa_rank,"-","supplementary_dataframe",".csv", sep="") 
write.csv(supplementary_dataframe, as.character(title_df), quote=FALSE, row.names=FALSE)
```